---
title: "Homework 6"
author: "Spencer Tibbitts"
date: "November 12, 2019"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#```{r, echo=F}

#```

## Introduction to the DataSet

The dataset 'brainsize.xlsx' contains data on a study of the
relationship between brain size and autism. The study randomly chose
30 autistic and 12 non-autistic boys who had received MRI scans as
toddlers.

```{r, echo=F}
library(xlsx)
brain <- read.csv('~/Downloads/brainsize.csv')
head(brain)
```

### Summarize the Brain Volumes of Each Group

Below is a numerical summary of the data by test group.

```{r, echo=F}

summary(brain)
```

### Visualizing the Data

It is easier to visualize this summary with a boxplot, as shown below.

```{r, echo=F}

# Create a boxplot to visualize the data.
boxplot(brain$Autistic, brain$Control,
horizontal=T,main="Brain Volume by
Group",names=c("Autistic","Control"),col=c("red","blue"))
```
Alternatively, we can use a histogram to visualize the differences
between the two groups.
```{r, echo=F}
# Now create a overlapping histogram to visualize the same data.
hist(brain$Autism, col=rgb(1,0,0,0.5),main="Brain Volume by
Group",xlab="Brain Volume")
hist(brain$Control, col=rgb(0,1,0,0.5),add=T)
box()
```

## Two Sample T-Test

We can perform a two sample t-test on the data to determine if there
truly is a difference between the two groups. Our hypothesi for this
experiment are as follows:

* $H_o$ - The mean brain volume for each group of boys is the same.
* $H_A$ - The mean brain volume for autistic boys is **greater** than
that of the control group.

```{r, echo=F}
t.test(brain,conf.level = 0.95)
```

My RStudio is refusing to compile.  If the p-value is below a certain level (less than 0.01) then this means we are able to reject the null hypothesis, and most likely accept the alternative hypothesis.  The interpretation for the confidence interval is that given two values, *a* and *b*, we are 95% confident that a random sample would fall between *a* and *b*.