---
title: "Homework 5"
author: "Spencer Tibbitts"
date: "10/22/2019"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```

## Example of Sampling Distribution
```{r load_height, include=F}
### Example
# a)
hgt<-read.csv("../Downloads/SienaMath-height.csv")
x_bar<-mean(hgt$height)
sd_hgth<-sd(hgt$height)
```

We have imported data provided on the height of the 34 Siena Math Majors.  From the data we can see that $\bar{x}\pm\sigma \to$ 
`r x_bar` 
$\pm$ 
`r sd_hgth`. 


Suppose we took a random sample of 3 people from the data, and wanted to compute the mean. 
```{r,include=T}
s<-sample(hgt$height,3)
mean(s)
```

Now let's do the same thing for 10 samples of 3 people at once.
```{r,include=T}
MeanXbar<-numeric(10)
for (i in 1:10){
  smp<-sample(hgt$height,3);
  MeanXbar[i]<-mean(smp)}
mean(MeanXbar)
```

Now lets do this with 100 samples, and plot the distribution of means as a histogram.
```{r part-d, echo=T}
MeanXbar=numeric(100)
for (i in 1:100){
  smp=sample(hgt$height,3);
  MeanXbar[i]=mean(smp)}
hist(MeanXbar,col="orange", xlab="Mean of Samples",main="")
```

From this random simulation, we found the mean and standard deviation of the sample means to be the following.
```{r, echo=T}
mean(MeanXbar)
sd(MeanXbar)
```
These results should be compared to the known values of $\mu$ and $\frac{\sigma}{\sqrt{n}}$ below.
```{r,echo=T}
mean(hgt$height)
sd(hgt$height)/sqrt(3)
```

There are $\frac{34!}{31!(3!)}=5984$ possible unique samples of size 3.  Let's select this many samples from the dataset and repeat the previous example.
```{r example-e, echo=T}
MeanXbar=numeric(5984)
for (i in 1:5984){
  smp=sample(hgt$height,3);
  MeanXbar[i]=mean(smp)}
hist(MeanXbar,col="green", xlab="Mean of Samples",main="")
mean(MeanXbar)
mean(hgt$height)
sd(MeanXbar)
sd(hgt$height)/sqrt(3)
```

## Three Simulations of the Sample Mean

Now let's run 3 more simulations, this time on the \it{Uniform}(0,1) distribution: first with a sample size of 1, then a sample size of 30, then a sample size of 1000.  It will become apparent that the greater the sample size, the more accurate a measurement of the mean is.
```{r example2, include=F}
xbar=numeric(1000)
for (i in 1:1000){
  x=runif(n=1,min=0,max=1);
  xbar[i]=mean(x)}
hist(xbar,col="yellow",main="sample size=1",xlab="sample mean Xbar")
mean(xbar)
sd(xbar)
```

```{r example2-b, include=F}
xbar2=numeric(1000)
for (i in 1:1000){
  x=runif(n=30,min=0,max=1);xbar2[i]=mean(x)}
hist(xbar2,col="red",main="sample size=30",xlab="sample mean Xbar")
mean(xbar2)
sd(xbar2)
xbar3=numeric(1000)
for (i in 1:1000){
  x=runif(n=1000,min=0,max=1);
  xbar3[i]=mean(x)}
hist(xbar3,col="green",main="sample size=1000",xlab="sample mean Xbar")
mean(xbar3)
sd(xbar3)
```

```{r,echo=T}
hist(xbar,col="yellow",main="sample size=1",xlab="sample mean Xbar")
hist(xbar2,col="red",main="sample size=30",xlab="sample mean Xbar")
hist(xbar3,col="green",main="sample size=1000",xlab="sample mean Xbar")
```