---
title: "Homework3-solution_TibbittsS"
author: "Spencer Tibbitts"
date: "10/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 3.6
#### If you roll a pair of fair dice, what is the probability of
*(a) getting a sum of 1?*
*(b) getting a sum of 5?*
*(c) getting a sum of 12?*

a) There is a 0% chance of rolling a sum of 1 with two dice - the only possible results are from 2-12.

b) There are 36 possible combinations for 2 dice.  A total of five can be rolled with the following pairs - (1,4), (4,1), (2,3), (3,2).
Therefore the probability of rolling a 5 is $\frac{4}{36} \to \frac{1}{9}$

c) There is only one possible combination to roll a 12 with - (6,6), therefore the probability is $\frac{1}{36}$.

### 3.9
#### In parts (a) and (b), identify whether the events are disjoint, independent, or neither (events cannot be both disjoint and independent).

(a) *You and a randomly selected student from your class both earn A's in this course.* 
**Independent**
(b) *You and your class study partner both earn A's in this course.*
**Neither**
(c) *If two events can occur at the same time, must they be dependent?*
**No**.  Say for instance, two car crashes occur at the same time of day - one in NYC and the other in Hong Kong.  Clearly, the two events are not related, therefore they are not dependent on one another.


### 3.16
#### The Behavioral Risk Factor Surveillance System (BRFSS) is an annual telephone survey designed to identify risk factors in the adult population and report emerging health trends. The following table displays the distribution of health status of respondents to this survey (excellent, very good, good, fair, poor) and whether or not they have health insurance.

*(a) Are being in excellent health and having health coverage mutually exclusive?*
No - the two definitely have some sort of relationship with one another. 

*(b) What is the probability that a randomly chosen individual has excellent health?*
$23.29\%$

*(c) What is the probability that a randomly chosen individual has excellent health given that he has health coverage?*
$20.99\%$

*(d) What is the probability that a randomly chosen individual has excellent health given that he doesn't have health coverage?*
$2.3\%$

*(e) Do having excellent health and having health coverage appear to be independent?*
While you can be in excellent health and not have health coverage, there is a relationship present between the two, so they are **not** independent.

### 3.20
#### A genetic test is used to determine if people have a predisposition for thrombosis, which is the formation of a blood clot inside a blood vessel that obstructs the flow of blood through the circulatory system. It is believed that 3% of people actually have this predisposition. The genetic test is 99% accurate if a person actually has the predisposition, meaning that the probability of a positive test result when a person actually has the predisposition is 0.99. The test is 98% accurate if a person does not have the predisposition. What is the probability that a randomly selected person who tests positive for the predisposition by the test actually has the predisposition?

The number of people who test positive are:

```{r, include=TRUE}
false_positive = 97*0.02
positive = 3*0.99
positive/(false_positive+positive)
```

### 3.30
#### Consider the following card game with a well-shuffled deck of cards. If you draw a red card, you win nothing. If you get a spade, you win \$5. For any club, you win \$10 plus an extra \$20 for the ace of clubs.

*(a) Create a probability model for the amount you win at this game. Also, find the expected winnings for a single game and the standard deviation of the winnings.*
P(\$0) = $\frac{1}{2}$
P(\$5) = $\frac{1}{4}$
P(\$10) = $\frac{3}{13}$
P(\$30) = $\frac{1}{52}$

```{r, echo=TRUE}
odds <- c(0.5,0.25,3/13,1/52)
results <- c(0,5,10,30)
N = 10000
samp <- sample(results,size=N,replace=T,prob=odds)
mean(samp)
```

*(b) What is the maximum amount you would be willing to pay to play this game? Explain your reasoning.*
At max, I'd be willing to bet \$3 because by the expected winnings of the game, I'd make approximately \$1 per game played.


### 3.38
#### The relative frequency table below displays the distribution of annual total personal income (in 2009 inflation-adjusted dollars) for a representative sample of 96,420,486 Americans. These data come from the American Community Survey for 2005-2009. This sample is comprised of 59% males and 41% females

*(a) Describe the distribution of total personal income.*

This distribution seems to be mostly normal with a longer right tail.

*(b) What is the probability that a randomly chosen US resident makes less than \$50,000 per year?* 

```{r, include=T}
.212+.183+.158+.047+.022
```

* (c) What is the probability that a randomly chosen US resident makes less than \$50,000 per year and is female? Note any assumptions you make. *
Assuming that there is no wage-gap and men and women are both paid equally in this survey, then this would be 
```{r, include=T}
0.622*0.41
```

*(d) The same data source indicates that 71.8% of females make less than $50,000 per year. Use this value to determine whether or not the assumption you made in part (c) is valid.*
If 71.8\% of females make less than \$50,000 per year, then this says the following percentage is the percentage of the population that is a female making less than \$50,000 per year:
```{r, include=T}
0.718*0.41
```
This is a higher percentage than expected, which means that men may be getting paid more than women on average (at least in this threshold).
