---
title: "Homework 4"
author: "Spencer Tibbitts"
date: "10/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=T}
download.file("http://www.openintro.org/stat/data/kobe.RData", destfile =
"kobe.RData")
load("kobe.RData")
head(kobe)
```
#### 1.) What does a streak length of 1 mean, i.e. how many hits and misses are in a streak of 1? What about a streak length of 0?
A streak length of 1 means that Kobe hit one shot, then missed the next one.  A streak length of 0 means that Kobe has just missed his second (or third, fourth, etc.) consecutive shot.

#### 2) Describe the distribution of Kobe’s streak lengths from the 2009 NBA finals. What was his typical streak length? How long was his longest streak of baskets?
```{r, include=T}
streaks <- calc_streak(kobe$basket)
typ_streak <- mean(streaks)
typ_streak # Mean
max_streak <- max(streaks)
max_streak # Max
```
An average streak is around 0.76, and the highest streak was 4 made shots in a row.

#### 3) In your simulation of flipping the unfair coin 100 times, how many flips came up heads?

```{r,include=T}
outcomes <- c("heads", "tails")
sim_unfair_coin <- sample(outcomes, size = 100, replace = TRUE, prob = c(0.2, 0.8))
table(sim_unfair_coin)
```


#### 4 What change needs to be made to the sample function so that it reflects a shooting percentage of 45%? Make this adjustment, then run a simulation to sample 133 shots. Assign the output of this simulation to a new object called sim_basket.

```{r,include=T}
outcomes <- c("H", "M") # Hit or miss
sim_basket <- sample(outcomes, size = 133, replace = TRUE, prob = c(0.45, 1-0.45))
table(sim_basket)
```

#### 5. Describe the distribution of streak lengths. What is the typical streak length for this simulated independent shooter with a 45% shooting percentage? How long is the player’s longest streak of baskets in 133 shots?
```{r,include=T}
streak_sim <- calc_streak(sim_basket)
mean(streak_sim) # Mean
max(streak_sim) # Max
```


#### 6. If you were to run the simulation of the independent shooter a second time, how would you expect its streak distribution to compare to the distribution from the question above? Exactly the same? Somewhat similar? Totally different? Explain your reasoning.

I would expect the outcome for a typical streak to be in the same ballpark, but not closer than within 0.1 of the previous result.  133 shots is a fairly small sample size to draw a definite conclusion from.  If the sample size was closer to 1,000 or 10,000 shots, then I would expect the results of the second simulation to be almost identical.

#### 7. How does Kobe Bryant’s distribution of streak lengths from page 2 compare to the distribution of streak lengths for the simulated shooter? Using this comparison, do you have evidence that the hot hand model fits Kobe’s shooting patterns? Explain.

```{r,include=T}
streak_sim
streaks
```

In the simulation of streaks, we know that the probability of every shot is a constant 45\% for certain, and it appears that there are only a few instances within the 133 shots where the shooter has multiple streaks in a row - and when they do, its almost always a streak of 1 followed by another streak of 1.  However, Kobe has multiple occurences of back to back to back streaks where you can see him get the hot hand.  The stretches of [1 3 3 1 1] and [3 2 0 3 0 1 3] definitely seem to indicate this.  Therefore there is evidence that Kobe fits the hot hand model.
